---
title: "Preparing The Data"
output: html_document
author: "Amir Zehavi, Adam Uziel and Itay Asraf"
id_Adam: "323817379"
id_itay: "322601907"
id_amir: "206655839"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preparing The Data
This file is the web scraping and preparation of the wiki covid_cases and covid_deaths in order to get a data-sets that are ready to data analysis.
After this code we started to answer the 3 questions.

```{r}
library(tidyverse)
library(rvest)
library(lubridate)
library(here)
library(glue)
library(countrycode)
library(dplyr, warn.conflicts = FALSE)
# Suppress summarise info
options(dplyr.summarise.inform = FALSE)
library(splitstackshape)
library(gridExtra)
library(data.table)

```

Taking the relevant tables, cases and death from Wikipedia.

```{r}
#The tables in the html are in class table=wikitable
Covid19_US<-read_html("https://en.wikipedia.org/wiki/Template:COVID-19_pandemic_data/United_States_medical_cases")

data_COVID19_US<-Covid19_US%>%
  html_nodes(".wikitable")%>%
  html_table(fil=TRUE)

#the 2 tables from wikipedia. The first-cases. Second-Died.
cases_us<-data.frame(data_COVID19_US[1])
death_cases<-data.frame(data_COVID19_US[2])
```

Matching The States to their areas, that he located in.

```{r}
#taking only the areas.
areas <- cases_us %>% 
  select(Date:Territories.3) %>%
  filter(Date == "Date"& West == "West") %>%
  head(1)

cases_US_left<- cases_us %>%
  select(Date:Territories.3) 
colnames(cases_US_left) <- c(cases_US_left[1,])



#taking the state and the area that each state belong there.
temp <- tibble(State = names(cases_US_left), area = as.vector(t(areas))) %>%
  filter(State!="Date")

```

Changing types of columns, and clean the irrelevant information
We chose to put 0 instead of null values in the both of tables, because we didn't get any information about the the day and we assumed that means no one died/got seek in thr particular date..

**We choose to put at null values the 0 value in the both of tables, because we didn't get any information about the the day and we couldn't know the cases/deaths.**

```{r}
#Because at the wikipedia tables there are columns, with values that we didnt want. Like:notes,refs,date(every month the date repeated on himself) and the total table
cases_US_left <- cases_US_left %>%
  filter(Date!="Date" & Date!="Total" & Date!= "Notes" & Date!= "Refs") %>%
  
   #pivot longer because we want in one c olumns to see the states and the another column the number of the cases
  pivot_longer(AK:VI, names_to = "State", values_to = "num_cases" ) %>%
  
  #correspond each state to his area.
  left_join(temp, by = "State")  %>%
  
  #dividing the date, because we want to change the format of this date.
  separate(Date, into = c("Day", "Month", "Year"), sep = "-") 

#if there are null values we will put 0
cases_US_left$num_cases<-suppressWarnings(as.integer(cases_US_left$num_cases))
cases_US_left$num_cases[is.na(cases_US_left$num_cases)] <- 0

#changing the type of the date
cases_US_left$Day <- as.integer(cases_US_left$Day)
cases_US_left$Year <- as.integer(cases_US_left$Year)
cases_US_left$Year <- cases_US_left$Year + 2000 #the year will be 2020,not 20

#changing from 3 letters month to numeric month
cases_US_left$Month<-match(cases_US_left$Month,month.abb)

#making date as year-month-day.
cases_US_left<-cases_US_left %>%
  mutate(Date = make_datetime(Year, Month, Day))

#change the tyope to date.
cases_US_left$Date <- as.Date(cases_US_left$Date)

#erase the columns of year,month,day of the date
drop <- c("Year","Month","Day")
cases_US_left = cases_US_left[,!(names(cases_US_left) %in% drop)]

```

```{r}
#erase the date column from the table
death_us<- death_cases %>%
  select(Date:Territories.3) 

colnames(death_us) <- c(death_us[1,])


#taking only the state and the area that the state in
temp2 <- tibble(State = names(death_us), area = as.vector(t(areas))) %>%
  filter(State!="Date")

death_us <- death_us %>%
  
  #Because at the wikipedia tables there are columns, with values that we didnt want. Like:notes,refs,date(every month the date repeated on himself) and the total table
  filter(Date!="Date" & Date!="Total" & Date!= "Notes" & Date!= "Refs") %>%
  
  #pivot longer because we want in one c olumns to see the states and the another column the number of the cases
  pivot_longer(AK:VI, names_to = "State", values_to = "num_deaths" ) %>%
  
  #correspond each state to his area.
  left_join(temp2, by = "State")  %>%
  
  #dividing the date, because we want to change the format of this date.
  separate(Date, into = c("Day", "Month", "Year"), sep = "-")


#change null values to 0 TRUE

death_us$num_deaths<-suppressWarnings(as.integer(death_us$num_deaths))
death_us$num_deaths[is.na(death_us$num_deaths)] <- 0

#changing the type of the date
death_us$Day <- as.integer(death_us$Day)
death_us$Year <- as.integer(death_us$Year)
death_us$Year <- death_us$Year + 2000

#changing from 3 letters month to numeric month
death_us$Month<-match(death_us$Month,month.abb)

#making date as year-month-day.  
death_us<-death_us %>% 
  mutate(Date = make_datetime(Year, Month, Day))

#change the type to date
death_us$Date <- as.Date(death_us$Date)

#erase the columns of year,month,day of the date
drop <- c("Year","Month","Day")
death_us = death_us[,!(names(death_us) %in% drop)]




```

Adding the full country names, in addition to their short names.

```{r, message=F, warning=F}

#change the short names of countries in usa to full name.
library(countrycode)
url = "https://raw.githubusercontent.com/vincentarelbundock/countrycode/master/data/custom_dictionaries/us_states.csv"

#at this, we have all the full names of usa countries.
state_dict = read.csv(url, stringsAsFactors=FALSE)
dea_country<-data.frame(countrycode(death_us$State, 'abbreviation', 'state', custom_dict=state_dict))

#Rename the column, more esthatic
dea_country<-dea_country%>%
  rename(full_name_state=countrycode.death_us.State...abbreviation....state...custom_dict...state_dict.)

#putting the full name state in our data frame. More esthatic
death_us_final<-death_us %>%
  mutate(full_name = dea_country$full_name_state)
write_csv(death_us_final, here::here("data", "Death_US_final.csv"))
write_csv(cases_US_left, here::here("data", "Cases_US_final.csv"))

```
